### 2024-04-17 paper/paper.qmd
Cleaned data and created model

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/deaths_months_prediction.rds"))
```

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Examining the number of e/E's as the number of words increase"
#| tbl-pos: H
#| echo: false
#| warning: false
#| eval: true

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

```{r}
#| label: fig-modelresultsfig
#| fig-cap: "The predicted number of e/E's in each line based on the number of words"
#| echo: false
#| warning: false
#| eval: true

library(marginaleffects)

plot_predictions(first_model, condition = "month_of_death") +
  labs(x = "Month",
       y = "Number of Actively Homeless") +
  theme_classic()
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-17 scripts/00-simulate_data.R

@@ -1,19 +1,43 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates Toronto Shelter System Flow data from 2018 to 2023.
+# Author: Irene Huynh
+# Date: 2 April 2024
+# Contact: irene.huynh@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+set.seed(123) # For reproducibility 
 
+n <- 504 # Number of observations/rows in the table
 
+# Dates for each row
+date_seq <- seq.Date(from=mdy("01-01-2018") , to=mdy("12-01-2023"), by="month")
 
+simulated_data <-
+  tibble(
+    # Use 1 through to 504 to represent each id
+    id = 1:504,
+    date = rep(x = date_seq, each = 7),
+    population_group = sample(c("Chronic", "Refugees", "Families", "Youth", "Single Adult", "Non-Refugees"), n, replace=TRUE),
+    # Randomly pick an integer, with replacement, n times
+    returned_from_housing = sample(1:200, n, replace=TRUE),
+    newly_identified = sample(1:2000, n, replace=TRUE),
+    actively_homeless = sample(1:11000, n, replace=TRUE),
+    age_under16 = sample(1:2000, n, replace=TRUE),
+    age_16_24 = sample(1:2000, n, replace=TRUE),
+    age_25_44 = sample(1:5000, n, replace=TRUE),
+    age_45_64 =sample(1:4000, n, replace=TRUE),
+    age_65over = sample(1:700, n, replace=TRUE),
+    gender_male = sample(1:7000, n, replace=TRUE),
+    gender_female = sample(1:5000, n, replace=TRUE),
+  )
+
+# Format the date as seen in the original data set
+simulated_data$date <- format(as.Date(simulated_data$date), "%b-%y")
+
+simulated_data

### 2024-04-17 scripts/01-download_data.R

@@ -1,26 +1,45 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the Toronto Shelter System Flow data and Deaths of People Experiencing Homelessness data from opendatatoronto.
+# Author: Irene Huynh
+# Date: 2 April 2024
+# Contact: irene.huynh@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
-
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+library(dplyr)
 
 
+#### Download shelter data ####
+toronto_shelters <-
+  # Each package is associated with a unique id  found in the "For 
+  # Developers" tab of the relevant page from Open Data Toronto
+  # https://open.toronto.ca/dataset/toronto-shelter-system-flow/
+  list_package_resources("ac77f532-f18b-427c-905c-4ae87ce69c93") |>
+  filter(name == "toronto-shelter-system-flow.csv") |>
+  # Having reduced the dataset to one row we can get the resource
+  get_resource()
 
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
-         
+#### Save shelter data ####
+write_csv(toronto_shelters, "data/raw_data/raw_toronto_shelters_data.csv") 
+
+
+#### Download homeless deaths data ####
+package <- show_package("a7ae08f3-c512-4a88-bb3c-ab40eca50c5e")
+package
+
+# Get all resources for this package
+resources <- list_package_resources("a7ae08f3-c512-4a88-bb3c-ab40eca50c5e")
+
+# Identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# Load the first datastore resource as a sample
[…]

### 2024-04-17 scripts/02-data_cleaning.R

@@ -1,44 +1,65 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw Toronto Shelter System Flow data and Deaths of People Experiencing Homelessness data from opendatatoronto.
+# Author: Irene Huynh
+# Date: 2 April 2024
+# Contact: irene.huynh@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+
 
 #### Workspace setup ####
 library(tidyverse)
+library(janitor)
+library(dplyr)
+library(arrow)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+#### Clean shelter data ####
+raw_toronto_shelters_data <- read_csv("data/raw_data/raw_toronto_shelters_data.csv")
 
-#### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+cleaned_toronto_shelters_data <-
+  raw_toronto_shelters_data |> 
[…]

### 2024-04-17 scripts/03-test_data.R

@@ -1,15 +1,61 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the cleaned Toronto Shelter System Flow data and Deaths of People Experiencing Homelessness data.
+# Author: Irene Huynh
+# Date: 2 April 2024
+# Contact: irene.huynh@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Test data ####
+
+#### Test homeless deaths data ####
+deaths_test <- read_csv("data/analysis_data/cleaned_homeless_deaths_data.csv")
+
+# Checking that there are exactly 78 entries/rows (12 months * 6 years + 6 months so far for 2023)
+nrow(deaths_test) == 78
+
+# Checking that the only years involved are from 2017 to 2023
+deaths_test$year_of_death %in% c(2017, 2018, 2019, 2020, 2021, 2022, 2023)
+
+# Checking that the only months involved are "January", "February", "March", "April", 
+# "May", "June", "July", "August", "September", "October", "November", and "December"
+deaths_test$month_of_death %in% c("January", "February", "March", "April", 
+                                   "May", "June", "July", "August", "September", 
+                                   "October", "November", "December")
+
+# Checking the classes of the columns
+deaths_test$year_of_death |> class() == "numeric"
+deaths_test$month_of_death |> class() == "character"
+deaths_test$count |> class() == "numeric"
+
+
+#### Test shelter data ####
+shelter_test <- read_csv("data/analysis_data/cleaned_toronto_shelters_data.csv")
+
+# Checking that there are exactly 564 entries/rows
+nrow(shelter_test) == 564
+
+# Checking that the only years involved are from 2018 to 2023
+substring(shelter_test$date, 5, 6) %in% c("18", "19", "20", "21", "22", "23", "24")
+
+# Checking that the only population groups involved are "Chronic", "Refugees", "Families", 
+# "Youth", "Single Adult", "Non-refugees", "Indigenous", and "All Population"
+shelter_test$population_group %in% c("Chronic", "Refugees", "Families", 
[…]

### 2024-04-17 scripts/04-model.R

@@ -1,37 +1,43 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the predicated month effect on homeless death.
+# Author: Irene Huynh
+# Date: 2 April 2024
+# Contact: irene.huynh@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(modelsummary)
+
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+cleaned_homeless_deaths_data <- read_parquet("data/analysis_data/cleaned_homeless_deaths_data.parquet")
+
 
 ### Model data ####
-first_model <-
+deaths_months_prediction <-
   stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
+    count ~ month_of_death,
+    data = cleaned_homeless_deaths_data,
+    family = poisson(link = "log"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
     seed = 853
   )
 
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  deaths_months_prediction,
+  file = "models/deaths_months_prediction.rds"
 )
 
 
+### View model summary ###
+modelsummary(
+  list(
+    "Month" = deaths_months_prediction
+  )
+)

### 2024-04-17 paper/paper.qmd
Wrote abstract, intro, and data; and added deaths simulation

@@
 title: "My title"
-subtitle: "My subtitle if needed"
 author: 
-  - First author
-  - Another author
+  - Irene Huynh
 thanks: "Code and data are available at: LINK."
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Toronto’s shelter system and homeless population has experienced a significant increase over recent years.
+Using Toronto’s shelter system data and deaths of people experiencing homelessness data from the City of Toronto's Open Data Portal, this paper analyzes when deaths most frequently occur in Toronto’s homeless population and the trends of different demographics that utilized shelters from 2018 to 2023.
+This paper studies the month of death of unhoused individuals, the age group that visits shelters the most, the trend of the average number of individuals who are actively homeless using shelters, the average number of people who return to shelters after acquiring housing, and the gap separating the average number of male and female shelter users from 2018 to 2023.
+This paper finds that deaths of unhoused individuals occur most frequently in January.
+Additionally, males and individuals aged 25 to 44 account for a significantly larger portion of shelter users.
+This study may be employed to inform policy makers of the demographic most at risk of homelessness and of where to target solutions for tackling the underlying issue of homelessness in Toronto."
 format: pdf
@@
 library(tidyverse)
-library(palmerpenguins)
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Toronto’s shelter system provides a temporary home to those who are in need of one.
+The system is essential for Toronto’s homeless population and refugees as the shelters uphold standards that support meals and laundry accommodations, mental health and harm reduction services, and a host of counselors to direct and aid individuals acquire permanent housing.
+Although the shelters provide a multitude of services and aid to secure permanent housing for the unhoused, there are still many who return to shelters after securing housing [@torontoshelters].
 
-The remainder of this paper is structured as follows. @sec-data....
+The affordable housing crisis is directly contributing to Toronto’s shelter availability crisis.
+As Toronto’s housing market has skyrocketed in the past few years, it is no wonder why Toronto’s shelters are receiving significantly more visitors.
+The City of Toronto states that, “One in three of Toronto’s homeowner and renter households (373,965 households or 32 per cent) experienced affordability issues,” [@cityToronto].
+According to CBC News, an estimated average of at least 273 individuals were turned away from shelters in June 2023, which was a record high for Toronto shelters [@cbcshelters].
+The amount of families seeking warm beds has been unmatched and as Toronto’s winters can be extremely frigid, shelter availability is becoming increasingly more critical.
+Moreover, as temperatures continue to reach record highs as a consequence of climate change, Toronto’s summers become increasingly more hazardous for unhoused individuals.
+
+Analyzing and studying Toronto’s shelter system data can reveal those within society that are struggling the most with securing housing.
+This paper analyzes the demographics that utilize Toronto’s shelter system and the month of deaths of unhoused individuals.
+Analyzing different age groups that utilize shelters are of particular interest to produce a predictive model that determines whether specific months impact the likelihood of death in Toronto’s homeless population.
+In addition, this paper is also concerned with the actively homeless population and different demographics, such as refugees and single adults, that stay in shelters.
+The main estimand of this study is the effect of months on deaths in individuals experiencing homelessness in Toronto.
+This paper finds that the largest demographic of shelter users are males and individuals aged 25 to 44, and the death of people experiencing homelessness occurs most frequently in January.
+
+The remainder of this paper is structured as follows: the Data section details the collection
+and processing of Toronto shelter system and deaths of unhoused individuals data.
+Additionally, the section includes discussion on the measurement of the original dataset and the variables examined in this study.
+The following Models section discusses the regression model used and the Results section displays the tables and figures that model the data.
+The Discussion section draws conclusions from those tables and figures and details the importance of my findings and weaknesses of the paper.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Sources
+The data utilized in this paper was retrieved from the City of Toronto's Open Data Portal [@opendatatoronto].
+The statistical programming language R was used to retrieve, clean, and process the data [@citeR].
+In particular, the following R packages were used: opendatatoronto [@opendatatoronto] and tidyverse [@tidyverse] for data acquisition, janitor [@janitor] and knitr [@knitr] for data cleaning and processing, and ggplot2 [@ggplot2] for creating figures.
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+## Toronto Shelter System Data
+The shelter system data acquired from the City of Toronto's Open Data Portal provided monthly statistics of Toronto's shelters from January 2018 to December 2023.
+For each month, the data set included information for the entire population, chronic visitors, refugees, families, youth, single adults, and non-refugees.
+The information of interest for each of those demographics are the number of individuals that returned to shelters after previously acquiring housing, those that are new to the shelter system, those that are actively homeless, those that identify as male and female, and the number of individuals in each age group in that demographic.
+The age groups consist of visitors under 16, 16 to 24, 25 to 44, 45 to 64, and 65 years of age or older.
+These
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+## Deaths of People Experiencing Homelessness Data
+The death count for individuals experiencing homelessness data acquired from the City of Toronto’s Open Data Portal provided monthly statistics of deaths from January 2017 to June 2023.
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
+## Measurement
+The City of Toronto’s Open Data Portal is Toronto's official source for city data relating to each of its divisions and agencies.
+Thus, the quality of the data used in this study is reliable; however, the Portal does state that there were limitations to the data sets.
+For the Toronto shelter system data, a limitation or possible avenue for measurement error for the data is that the data excludes individuals who do not sleep in shelters or who use other resources for shelter, such as public infrastructure and buildings open all hours  [@torontosheltersystemflow].
+Additionally, for the deaths of people experiencing homelessness data, the data only includes reported deaths [@homelessdeathsdata].
+Therefore, the number of deaths could be much higher as unreported deaths were excluded from the dataset.
 
@@
 
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-04-17 scripts/00-simulate_data.R

@@ -10,7 +10,7 @@
 library(tidyverse)
 
 
-#### Simulate data ####
+#### Simulate shelter data ####
 set.seed(123) # For reproducibility 
 
 n <- 504 # Number of observations/rows in the table
@@ -18,7 +18,7 @@ n <- 504 # Number of observations/rows in the table
 # Dates for each row
 date_seq <- seq.Date(from=mdy("01-01-2018") , to=mdy("12-01-2023"), by="month")
 
-simulated_data <-
+simulated_shelter_data <-
   tibble(
     # Use 1 through to 504 to represent each id
     id = 1:504,
@@ -38,6 +38,21 @@ simulated_data <-
   )
 
 # Format the date as seen in the original data set
-simulated_data$date <- format(as.Date(simulated_data$date), "%b-%y")
+simulated_shelter_data$date <- format(as.Date(simulated_shelter_data$date), "%b-%y")
 
-simulated_data
+simulated_shelter_data
+
+
+### Simulate homeless deaths data ###
+n <- 84 # 7 years * 12 months in each year
+
+simulated_deaths_data <- 
+  tibble(
+    year = rep(x = c(2017, 2018, 2019, 2020, 2021, 2022, 2023), each = 12),
+    month = rep(x = c("January", "February", "March", "April", 
+                     "May", "June", "July", "August", "September", 
+                     "October", "November", "December"), each = 1, times = 7),
+    count = sample(1:50, n, replace=TRUE)
+    )
+
+simulated_deaths_data

### 2024-04-17 paper/paper.qmd
Added tables and figures and wrote results and model section

@@
 This study may be employed to inform policy makers of the demographic most at risk of homelessness and of where to target solutions for tackling the underlying issue of homelessness in Toronto."
-format: pdf
+format: 
+  pdf:
+    fig-pos: 'H'
+header-includes:
+  - \usepackage{float} 
+  - \floatplacement{table}{H}
 number-sections: true
@@
 library(tidyverse)
+library(dplyr)
+
+cleaned_toronto_shelters_data <- read_csv("../data/analysis_data/cleaned_toronto_shelters_data.csv")
 ```
@@
 The statistical programming language R was used to retrieve, clean, and process the data [@citeR].
-In particular, the following R packages were used: opendatatoronto [@opendatatoronto] and tidyverse [@tidyverse] for data acquisition, janitor [@janitor] and knitr [@knitr] for data cleaning and processing, and ggplot2 [@ggplot2] for creating figures.
+In particular, the following R packages were used: `opendatatoronto` [@opendatatoronto], `tidyverse` [@tidyverse], and `dplyr` [@dplyr] for data acquisition; `janitor` [@janitor] and `arrow` [@arrow] for data cleaning and processing; `rstanarm` [@rstanarm] and `marginaleffects` [@marginaleffects] for modeling; and `ggplot2` [@ggplot2] for creating figures.
 
@@
 The age groups consist of visitors under 16, 16 to 24, 25 to 44, 45 to 64, and 65 years of age or older.
-These
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of my modelling strategy is to utilize the Poisson distribution to build a Poisson regression model as the data is concerned with count data and $\lambda$ distributes probabilities over non-negative integers.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the number of deaths and the explanatory variable as the month. I am interested in how the number of deaths differs by month.
 
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i|\lambda_i &\sim \mbox{Poisson}(\lambda_i) \\
+\mbox{log}(\lambda_i) &= \beta_0 + \beta_1 \times \mbox{Month}_i\\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+I run the model in R [@citeR] using the `rstanarm` package of @rstanarm. I use the default priors from `rstanarm`.
 
@@
 
-Our results are summarized in @tbl-modelresults.
-
+## Relationship Between Deaths of Homeless Individuals and Months
 ```{r}
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Examining the number of e/E's as the number of words increase"
+#| tbl-cap: "Examining the number of deaths of homeless individuals occuring in each month"
 #| tbl-pos: H
@@
 
+## Relationship Between Deaths of Homeless Individuals and Months
+@tbl-modelresultsslopes helps with interpreting the model results in @tbl-modelresults.
+@tbl-modelresultsslopes suggests that in our dataset, compared to April, August tends to have around three additional deaths, December tends to have around two additional deaths, February tends to have one additional death, and so on.
+Most notably, compared to April, January tends to have five additional deaths–the most additional deaths out of all the months; July tends to have around four additional deaths–the second most additional deaths out of all the months; and March tends to have one less death–the least amount of additional deaths out of all the months.
+
 ```{r}
-#| label: fig-modelresultsfig
-#| fig-cap: "The predicted number of e/E's in each line based on the number of words"
+#| label: tbl-modelresultsslopes
+#| tbl-cap: "The estimated difference in the number of deaths of homeless indivuals in each month"
+#| tbl-pos: H
 #| echo: false
@@
 
+library(knitr)
 library(marginaleffects)
 
-plot_predictions(first_model, condition = "month_of_death") +
-  labs(x = "Month",
-       y = "Number of Actively Homeless") +
-  theme_classic()
+slopes(first_model) |>
+  select(contrast, estimate) |>
+  unique() |> 
+  kable(
+    col.names = c("Compare month", "Estimate"),
+    digits = 2, booktabs = TRUE, linesep = ""
+  )
+```
+
+```{r}
+#| label: fig-modelresultsfig
+#| fig-cap: "The predicted number deaths of homeless individuals based on the month"
+#| echo: false
+#| warning: false
+#| eval: true
+
+#library(marginaleffects)
+
+#plot_predictions(first_model, condition = "month_of_death") +
+#  labs(x = "Month",
+#       y = "Number of Deaths") +
+#  theme_classic()
+```
+
+## Average Number of Indivduals for each Age Group
+
+@fig-agegroupavg demonstrates that individuals age 25 to 44 enter shelters the most on average, followed by individuals age 45 to 64, then those under 16 and those age 16 to 24.
+Additionally, individuals 65 and over enter shelters the least on average.
+This result is expected as the cost of living in Toronto has risen significantly in recent years and individuals age 25 to 44 must have a large percentage of their income allotted to rent [@cbcliving].
+This quota may be difficult to reach every month as the price of other necessities have also risen, such as grocery bills.
+Individuals over 65 may not choose to stay in shelters as they have other options, including retirement homes and elder care centres.
+```{r}
+#| label: fig-agegroupavg
+#| fig-cap: "The average of all shelter visitors in each age group over 6 years from 2018 to 2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+age_group_data <-  cleaned_toronto_shelters_data|>
+    filter(population_group == "All Population") |>
+  summarise(
+    # Divide by (6*12) since it is 6 years (2018-2023) with 12 months each
+    "Under 16" = sum(age_under16, na.rm = TRUE)/(6*12), 
+    "16-24" = sum(age_16_24, na.rm = TRUE)/(6*12),  
+    "25-44" = sum(age_25_44, na.rm = TRUE)/(6*12),   
+    "45-64" = sum(age_45_64, na.rm = TRUE)/(6*12), 
+    "65+" = sum(age_65over, na.rm = TRUE)/(6*12)
+  ) |>
+  pivot_longer(cols = everything(), names_to = "age_group", values_to = "average")
+
+# Create the bar graph
+ggplot(age_group_data, aes(x = age_group, y = average, fill = age_group)) +
+  geom_bar(stat = "identity") +
+  labs(x = "Age Group",
+       y = "Average")
+```
+
+## Average Number of Actively Homeless Indivduals
+
+@fig-activelyhomeless demonstrates that the average number of actively homeless individuals increased from 2018 to 2019, significantly decreased from 2019 to 2020, remained almost unchanged from 2020 to 2021, experienced a sharp increase from 2021 to 2022, and increased more from 2022 to 2023.
+Additionally, the average in 2023 reached a high of almost 10500 actively homeless individuals.
+The housing crisis and spike in the cost of living since the COVID-19 pandemic is a notable contributing factor to the increase in homelessness and shelter usage after 2021.
+The significant decrease in the average number of actively homeless individuals using shelters from 2019 to 2021 can also be attributed to the pandemic: many public areas were restricted or fully closed for physical distancing purposes.
+```{r}
+#| label: fig-activelyhomeless
+#| fig-cap: "The average number of actively homeless people from 2018 to 2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+avgs_actively_homeless <- rep(NA, times=6) # 6 years (2018 to 2023)
+year = 18
+for (i in 1:6){
+  each_year_actively_homeless <- cleaned_toronto_shelters_data |>
+    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
+    select(actively_homeless)
+
+  avgs_actively_homeless[i] <- sum(each_year_actively_homeless)/12 # 12 months in a year
+
+  year = year + 1
+}
+
+avgs_actively_homeless_data <-
+  tibble(
+    date = 2018:2023,
+    avg = avgs_actively_homeless
+  )
+
+ggplot(avgs_actively_homeless_data, aes(x = date, y = avg, fill = date)) +
+  geom_line(stat = "identity") +
+  labs(x = "Year",
+       y = "Average")
+```
+
+## Average Number of Shelter Returners
+
+@fig-returnedfromhousingavg illustrates that the average number of individuals who return to shelters each year increased each year from 2018 to 2021, then the average began to decrease from 2021 to 2022 and significantly declined from 2022 to 2023.
+In addition, less than 60 individuals returned to a shelter in 2023.
+This outcome is surprising as it was expected to find that after acquiring housing, many would return to shelters due to the rise of rental prices and the overall housing and cost of living crisis in recent years.
+Additionally, @fig-activelyhomeless demonstrates the significant increase in homeless individuals.
+A reason that can explain why many homeless people who acquire housing and become homeless again but do not return to shelters is that more individuals are experiencing homelessness; thus, shelter capacity is quickly filling up and many returning visitors are being turned away [@cbcshelters].
+
+```{r}
+#| label: fig-returnedfromhousingavg
+#| fig-cap: "The average of number of visitors who returned to shelters from 2018 to 2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+avgs_returned <- rep(NA, times=6) # 6 years (2018 to 2023)
+year = 18
+for (i in 1:6){
+  each_returned_from_housing <- cleaned_toronto_shelters_data |>
+    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
+    select(returned_from_housing)
+
+  avgs_returned[i] <- sum(each_returned_from_housing)/12 # 12 months in a year
+
+  year = year + 1
+}
+
+avgs_returned_data <-
+  tibble(
+    date = 2018:2023,
+    avg = avgs_returned
+  )
+
+ggplot(avgs_returned_data, aes(x = date, y = avg, fill = date)) +
+  geom_line(stat = "identity") +
+  labs(x = "Year",
+       y = "Average")
+```
+
+## Average Number of Males and Females
+
+There is an average of over 5000 male individuals in shelters each year (blue); whereas for female individuals (red), there is an average of less than 4000.
+@fig-malefemale illustrates how every year, the average number of male shelter users is significantly higher than that of female users.
+Additionally, from 2020 to 2021, the average number of male shelter users increased while the average number of female shelter users decreased.
+In regards to other annual intervals, the average for males and females both increase from 2018 to 2019, decrease from 2019 to 2020, and increase from 2021 to 2023.
+
+```{r}
+#| label: fig-malefemale
+#| fig-cap: "The average of male and female shelter visitors from 2018 to 2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+avgs_male <- rep(NA, times=6) # 6 years (2018 to 2023)
+avgs_female <- rep(NA, times=6)
+year = 18
+for (i in 1:6){
+  each_year_male <- cleaned_toronto_shelters_data |>
+    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
+    select(gender_male)
+
+  each_year_female <- cleaned_toronto_shelters_data |>
+    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
+    select(gender_female)
+
+  avgs_male[i] <- sum(each_year_male)/12 # 12 months in a year
+  avgs_female[i] <- sum(each_year_female)/12
+
+  year = year + 1
+}
+
+avgs_gender_data <-
+  tibble(
+    date = 2018:2023,
+    avg_males = avgs_male,
+    avg_females = avgs_female
+  )
+
+ggplot(avgs_gender_data, aes(x=date)) +
+  geom_line(aes(y = avg_males), color="blue") +
+  geom_line(aes(y = avg_females), color="red")+
+  labs(x = "Year",
+       y = "Average")
 ```
@@
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
+## First discussion point
 

### 2024-04-17 scripts/02-data_cleaning.R

@@ -59,6 +59,14 @@ cleaned_homeless_deaths_data <-
   filter(month_of_death != "Unknown") |>
   select(year_of_death, month_of_death, count)
 
+months = c("January", "February", "March", "April", "May", "June", "July", "August", "September",
+           "October", "November", "December")
+
+# Rename all months to only the first three letters of the month
+for (i in 1:12){
+  cleaned_homeless_deaths_data$month_of_death[cleaned_homeless_deaths_data$month_of_death==months[i]] <- substring(months[i], 1, 3)
+}
+
 
 #### Save homeless deaths data ####
 write_csv(cleaned_homeless_deaths_data, "data/analysis_data/cleaned_homeless_deaths_data.csv")

### 2024-04-17 scripts/03-test_data.R

@@ -21,9 +21,9 @@ deaths_test$year_of_death %in% c(2017, 2018, 2019, 2020, 2021, 2022, 2023)
 
 # Checking that the only months involved are "January", "February", "March", "April", 
 # "May", "June", "July", "August", "September", "October", "November", and "December"
-deaths_test$month_of_death %in% c("January", "February", "March", "April", 
-                                   "May", "June", "July", "August", "September", 
-                                   "October", "November", "December")
+deaths_test$month_of_death %in% c("Jan", "Feb", "Mar", "Apr", 
+                                   "May", "Jun", "Jul", "Aug", "Sep", 
+                                   "Oct", "Nov", "Dec")
 
 # Checking the classes of the columns
 deaths_test$year_of_death |> class() == "numeric"

### 2024-04-17 paper/paper.qmd
Wrote datasheet, added references

@@
 
-# Data {#sec-data}
+# Data
 
@@
 ## Relationship Between Deaths of Homeless Individuals and Months
+
 @tbl-modelresultsslopes helps with interpreting the model results in @tbl-modelresults.
@@
 Individuals over 65 may not choose to stay in shelters as they have other options, including retirement homes and elder care centres.
+
 ```{r}
@@
 The significant decrease in the average number of actively homeless individuals using shelters from 2019 to 2021 can also be attributed to the pandemic: many public areas were restricted or fully closed for physical distancing purposes.
+
 ```{r}
@@
 
-## First discussion point
+## Findings
 

### 2024-04-17 paper/datasheet.qmd
Wrote discussion, README, and cleaned up files

---
title: "Datasheet for 'Toronto Shelter System Flow Data'"
author: 
  - Irene Huynh
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the demographics of Toronto's shelter system occupants from 2018 to 2023.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Irene Huynh, while working as a student at the University of Toronto.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - No direct funding or grants were provided for the creation of the dataset.
4. *Any other comments?*
    - No.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each row consists of monthly statistics of Toronto's shelters. These monthly statistics include, count data for various demographics of people using the shelters.
2. *How many instances are there in total (of each type, if appropriate)?*
	- There are exactly 564 instances.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is a contains all possible instances of individuals who used a Toronto shelter from 2018 to 2023.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each row consists of monthly statistics of Toronto's shelters. These monthly statistics include, count data for the population groups (entire population, chronic visitors, refugees, families, youth, single adults, and non-refugees), the number of individuals that returned to shelters after previously acquiring housing, those that are new to the shelter system, those that are actively homeless, those that identify as male and female, and the number of individuals in each age group in that demographic. The age groups consist of visitors under 16, 16 to 24, 25 to 44, 45 to 64, and 65 years of age or older.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, there is a unique ID number associated with each row.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- No.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, through the unique ID.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Self-contained.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Yes, population groups include chronic visitors, refugees, families, youth, single adults, and non-refugees. Additionally, there are columns separating gender (male and female) and age groups (visitors under 16, 16 to 24, 25 to 44, 45 to 64, and 65 years of age or older).
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No.
16. *Any other comments?*
	- No.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data were gathered from the City of Toronto's Open Data Portal.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Scraping and parsing using R.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is not a sample.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Irene Huynh, a student at the University of Toronto.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected over the span of 6 years (2018 to 2023).
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was collected directly from the City of Toronto's Open Data Portal.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- No.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- No.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Consent was not obtained.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.
12. *Any other comments?*
	- No.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes cleaning of the data was done.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, and the raw data is available at the City of Toronto's Open Data Portal at https://open.toronto.ca/dataset/toronto-shelter-system-flow/
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- R was used.
4. *Any other comments?*
	- No.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- No.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.
3. *What (other) tasks could the dataset be used for?*
	- Determining which population group comprises most of Toronto's homeless population.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- No.
6. *Any other comments?*
	- No.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset is available through GitHub.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- GitHub for now.
3. *When will the dataset be distributed?*
	- The dataset is available now.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No. MIT license.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- None that are known.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- None that are known.
7. *Any other comments?*
	- No.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Irene Huynh
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- irene.huynh@mail.utoronto.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No, the dataset is just updated.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- No.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- No.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- No.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Pull request on GitHub.
8. *Any other comments?*
	- No.

\newpage

# References

### 2024-04-17 paper/paper.qmd
Wrote discussion, README, and cleaned up files

@@
 ---
-title: "My title"
+title: "An Analysis of the Demographics of Toronto's Homeless Population and the Relationship Between the Number Deaths of Homeless Individuals Based on the Month"
 author: 
@@
 This paper studies the month of death of unhoused individuals, the age group that visits shelters the most, the trend of the average number of individuals who are actively homeless using shelters, the average number of people who return to shelters after acquiring housing, and the gap separating the average number of male and female shelter users from 2018 to 2023.
-This paper finds that deaths of unhoused individuals occur most frequently in January.
+This paper finds that deaths of unhoused individuals occur most frequently in January and July.
 Additionally, males and individuals aged 25 to 44 account for a significantly larger portion of shelter users.
@@
 Toronto’s shelter system provides a temporary home to those who are in need of one.
-The system is essential for Toronto’s homeless population and refugees as the shelters uphold standards that support meals and laundry accommodations, mental health and harm reduction services, and a host of counselors to direct and aid individuals acquire permanent housing.
+The system is essential for Toronto’s homeless population and refugees as the shelters uphold standards that support meals and laundry accommodations, mental health and harm reduction services, and a host of counselors to direct and help individuals acquire permanent housing.
 Although the shelters provide a multitude of services and aid to secure permanent housing for the unhoused, there are still many who return to shelters after securing housing [@torontoshelters].
@@
 The affordable housing crisis is directly contributing to Toronto’s shelter availability crisis.
-As Toronto’s housing market has skyrocketed in the past few years, it is no wonder why Toronto’s shelters are receiving significantly more visitors.
-The City of Toronto states that, “One in three of Toronto’s homeowner and renter households (373,965 households or 32 per cent) experienced affordability issues,” [@cityToronto].
+As Toronto’s housing market has skyrocketed in the past few years, it is no surprise that Toronto’s shelters are receiving significantly more visitors.
+The City of Toronto states that as of 2023, “One in three of Toronto’s homeowner and renter households (373,965 households or 32 per cent) experienced affordability issues,” [@cityToronto].
 According to CBC News, an estimated average of at least 273 individuals were turned away from shelters in June 2023, which was a record high for Toronto shelters [@cbcshelters].
@@
 This paper analyzes the demographics that utilize Toronto’s shelter system and the month of deaths of unhoused individuals.
-Analyzing different age groups that utilize shelters are of particular interest to produce a predictive model that determines whether specific months impact the likelihood of death in Toronto’s homeless population.
+Analyzing the deaths of individuals experiencing homelessness is of particular interest to produce a predictive model that determines whether specific months impact the likelihood of death in Toronto’s homeless population.
 In addition, this paper is also concerned with the actively homeless population and different demographics, such as refugees and single adults, that stay in shelters.
 The main estimand of this study is the effect of months on deaths in individuals experiencing homelessness in Toronto.
-This paper finds that the largest demographic of shelter users are males and individuals aged 25 to 44, and the death of people experiencing homelessness occurs most frequently in January.
+This paper finds that the largest demographic of shelter users are males and individuals aged 25 to 44, and the death of people experiencing homelessness occurs most frequently in January and July.
 
@@
 ## Sources
+
 The data utilized in this paper was retrieved from the City of Toronto's Open Data Portal [@opendatatoronto].
 The statistical programming language R was used to retrieve, clean, and process the data [@citeR].
-In particular, the following R packages were used: `opendatatoronto` [@opendatatoronto], `tidyverse` [@tidyverse], and `dplyr` [@dplyr] for data acquisition; `janitor` [@janitor] and `arrow` [@arrow] for data cleaning and processing; `rstanarm` [@rstanarm] and `marginaleffects` [@marginaleffects] for modeling; and `ggplot2` [@ggplot2] for creating figures.
+In particular, the following R packages were used: `opendatatoronto` [@opendatatoronto], `tidyverse` [@tidyverse], and `dplyr` [@dplyr] for data acquisition; `janitor` [@janitor] and `arrow` [@arrow] for data cleaning and processing; `rstanarm` [@rstanarm], `marginaleffects` [@marginaleffects], and `knitr` [@knitr] for modeling; and `ggplot2` [@ggplot2] for creating figures.
 
 ## Toronto Shelter System Data
+
 The shelter system data acquired from the City of Toronto's Open Data Portal provided monthly statistics of Toronto's shelters from January 2018 to December 2023.
@@
 The age groups consist of visitors under 16, 16 to 24, 25 to 44, 45 to 64, and 65 years of age or older.
+## Deaths of People Experiencing Homelessness Data
 
-## Deaths of People Experiencing Homelessness Data
 The death count for individuals experiencing homelessness data acquired from the City of Toronto’s Open Data Portal provided monthly statistics of deaths from January 2017 to June 2023.
@@
 ## Measurement
+
 The City of Toronto’s Open Data Portal is Toronto's official source for city data relating to each of its divisions and agencies.
@@
 
-## Weaknesses and next steps
+This study utilized a Poisson regression model to predict the number of deaths of homeless individuals based on the month.
+The analysis conducted in this study found that the most deaths occurred in January, the second most deaths occurred in July, and the least deaths occurred in March.
+These findings are expected considering that January can consist of Toronto’s most frigid, cold weather, causing homeless individuals who cannot secure shelter to suffer in the extremely cold temperatures.
+July consists of the other extreme weather case: severe heat waves, which can be fatal.
+This study also found that the most prevalent age group that uses shelters is the 25 to 44 age range and the number of males that use shelters significantly surpass the number of females.
+Additionally, the number of actively homeless people in Toronto is trending upwards as of 2021.
+
+## Importance
+
+Housing is becoming increasingly difficult to acquire in Toronto, leaving thousands unhoused as temperatures reach their extremes in January and July.
+Many factors can be to blame, including the cost of living crisis, the housing crisis, and the COVID-19 pandemic.
+According to a new report, “The cost of living in Toronto went up nearly 20 percent between 2017 and 2022–but that’s less than half the increase seen in the city’s housing prices over the last five years,” [@cp24].
+As the cost of necessities soar, meeting rent payments becomes exceedingly difficult to accomplish, causing more and more people to find themselves unhoused.
+Homeless individuals are in vulnerable positions, but are often stigmatized as they face discrimination as a direct result of their housing status.
+Professionals and businesses are less willing to offer their services, leading to higher rates of poor physical and mental health.
+According to @stigma, there is evidence that “homeless individuals are not perceived as fully human,” making it tremendously distressing when searching for employment, healthcare, and sympathy.
+As much of society vilifies the unhoused, it is important to identify ways to provide relief for them.
+This paper reveals that deaths in Toronto’s homeless population occur most often in the coldest and warmest months of the year.
+With this knowledge, the City of Toronto can establish more temporary shelters during January and July and if that is not feasible, funding can be directed to establishing warming stations that can quickly be erected in January and to establishing cooling stations in July.
+
+## Weaknesses and Future Proceedings
+
+Although this paper used data from a reliable source, there were some limitations to the collection of the data which became potential sources for measurement error.
+Such limitations included the exclusion of unreported deaths of homeless individuals and shelters that are not part of the City of Toronto’s database.
+Moreover, the data used did not include the cause of death of individuals, leading to the speculation that the weather is solely responsible for the deaths; however, there are many other possibilities, such as underlying health conditions or drug overdoses.
+In future studies of homelessness and shelter usage in Toronto, an analysis of the cause of death of homeless individuals would provide more insight into where to target relief efforts.
 

### 2024-04-17 scripts/00-simulate_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulates Toronto Shelter System Flow data from 2018 to 2023.
+# Purpose: Simulates Toronto Shelter System Flow data from 2018 to 2023 and Toronto homeless deaths data.
 # Author: Irene Huynh
 # Date: 2 April 2024
 # Contact: irene.huynh@mail.utoronto.ca

### 2024-04-17 scripts/04-model.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Models the predicated month effect on homeless death.
+# Purpose: Models the predicated number of deaths of homeless individuals based on the month.
 # Author: Irene Huynh
 # Date: 2 April 2024
 # Contact: irene.huynh@mail.utoronto.ca

### 2024-04-17 paper/datasheet.qmd
Added git repo to paper and datasheet

@@
   - Irene Huynh
-thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+thanks: "Code and data are available at: https://github.com/IreneeeH/Homelessness_in_Toronto.git"
+date: today
+date-format: long
 format: pdf
@@
 Please provide a description.*
-	- Each row consists of monthly statistics of Toronto's shelters. These monthly statistics include, count data for various demographics of people using the shelters.
+	- Each row consists of monthly statistics of Toronto's shelters. These monthly statistics include count data for various demographics of people using the shelters.
 2. *How many instances are there in total (of each type, if appropriate)?*

### 2024-04-17 paper/paper.qmd
Added git repo to paper and datasheet

@@
   - Irene Huynh
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/IreneeeH/Homelessness_in_Toronto.git"
 date: today

### 2024-04-17 paper/paper.qmd
Edited title and README

--- 2024-04-17 paper/paper.qmd
@@
 ---
-title: "An Analysis of the Demographics of Toronto's Homeless Population and the Relationship Between the Number Deaths of Homeless Individuals Based on the Month"
+title: "An Analysis of the Demographics of Toronto's Homeless Population and the Relationship Between the Number of Deaths of Homeless Individuals Based on the Month"
 author: 
   - Irene Huynh
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/IreneeeH/Homelessness_in_Toronto.git"
 date: today
@@
 #| label: fig-modelresultsfig
-#| fig-cap: "The predicted number deaths of homeless individuals based on the month"
+#| fig-cap: "The predicted number of deaths of homeless individuals based on the month"
 #| echo: false

### 2024-04-17 paper/paper.qmd
Edited abstract

@@
 abstract: "Toronto’s shelter system and homeless population has experienced a significant increase over recent years.
-Using Toronto’s shelter system data and deaths of people experiencing homelessness data from the City of Toronto's Open Data Portal, this paper analyzes when deaths most frequently occur in Toronto’s homeless population and the trends of different demographics that utilized shelters from 2018 to 2023.
-This paper studies the month of death of unhoused individuals, the age group that visits shelters the most, the trend of the average number of individuals who are actively homeless using shelters, the average number of people who return to shelters after acquiring housing, and the gap separating the average number of male and female shelter users from 2018 to 2023.
+Using Toronto’s shelter system data and deaths of people experiencing homelessness data from the City of Toronto's Open Data Portal, this paper analyzes the month of death of unhoused individuals, the age group that visits shelters the most, the trend of the average number of individuals who are actively homeless using shelters, the average number of people who return to shelters after acquiring housing, and the gap separating the average number of male and female shelter users from 2018 to 2023.
 This paper finds that deaths of unhoused individuals occur most frequently in January and July.
